---
title: "Energy consumption and weather"
author: "Michael Tyler"
date: "15 February 2018"
subtitle: "An interactive app for examining relationships between different weather variables and aggregated electricity consumption of samples of London households."
output: ioslides_presentation
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Inputs and outputs

```{r echo=FALSE, message=FALSE}
library(lubridate)
library(plotly)

df <- read.csv('kWh_weather.csv')
df$interval <- ymd_hms(df$interval)
```

The inputs consist of four radio buttons and a slider.

- The radio buttons are used to select a weather variable.
- The slider is used to select a sample size between 1 and 100.
- When the user clicks the "Create plots" button, a sample of households of the selected size is taken from the dataset, and the energy consumption is summed over these households.  A new sample is taken each time the button is clicked.

The app then creates two plots.

- The first is a plot of the aggregated consumption against the selected weather variable, fitted with a linear model.
- The second plot shows the energy consumption profile for a typical winter day, aggregated over the chosen sample.

```{r echo=FALSE, message=FALSE}
df <- read.csv('kWh_weather.csv')
df$interval <- ymd_hms(df$interval)
tempdf <- cbind(df[, c('interval', 'temp', 'wspd'), drop = FALSE], kWh = rowSums(df[, -(1:6)]))
df_13Feb <- df[as.character(as.Date(df$interval)) == '2013-02-13', ]
tempdf_13Feb <- tempdf[as.character(as.Date(tempdf$interval)) == '2013-02-13', ]
```

## First plot

The correlation becomes more consistent as the sample size increases.  Here is an example with temperature - 3 samples of size 1 produce 3 very different results, but as the sample size increases to 100 the relationship converges to the lower graph:

```{r echo=FALSE, message=FALSE, warning=FALSE}
p1 <- plot_ly(df, x = ~temp, y = ~N0191, type = 'scatter', mode = 'markers',
              marker = list(size = 2), height = 150) %>%
  layout(xaxis = list(title = ''))
mod1 <- lm(N0191 ~ temp, df)
p1mod <- add_lines(p1, x = ~temp, y ~ fitted(mod1))

p2 <- plot_ly(df, x = ~temp, y = ~N0177, type = 'scatter', mode = 'markers',
              marker = list(size = 2), height = 150) %>%
  layout(xaxis = list(title = ''))
mod2 <- lm(N0177 ~ temp, df)
p2mod <- add_lines(p2, x = ~temp, y ~ fitted(mod2))

p3 <- plot_ly(df, x = ~temp, y = ~N0240, type = 'scatter', mode = 'markers',
              marker = list(size = 2), height = 150) %>%
  layout(xaxis = list(title = ''))
mod3 <- lm(N0240 ~ temp, df)
p3mod <- add_lines(p3, x = ~temp, y ~ fitted(mod3))

p <- subplot(p1mod, p2mod, p3mod, nrows = 1, shareY = TRUE) %>%
  layout(yaxis = list(title = 'kWh', range = c(0, 0.75)))

p4 <- plot_ly(tempdf, x = ~temp, y = ~kWh, type = 'scatter', mode = 'markers', height = 350) %>%
  layout(xaxis = list(title = 'Temperature'))
mod <- lm(kWh ~ temp, tempdf)
p4mod <- add_lines(p4, x = ~temp, y ~ fitted(mod))

subplot(p, p4mod, nrows = 2, heights = c(0.3, 0.7), titleX = TRUE, titleY = TRUE) %>%
  layout(showlegend = FALSE)
```

## Summary table

Beneath the first plot is a summary table for the linear model, along with a sentence stating whether or not the correlation is statistically significant.  Here is an example for the lower graph on the previous slide.  The linear model object is `mod`.

```{r message=FALSE, tidy=TRUE, tidy.opts=list(width.cutoff=60)}
cat(if(summary(mod)$coeff[2,4] < 0.05) {
      paste0('p = ', signif(summary(mod)$coeff[2,4], 2), ', so there', ' is a statistically significant correlation\nbetween temp', ' and kWh in this sample.')
    } else {
      paste0('p = ', signif(summary(mod)$coeff[2,4], 2), ', so there', ' is no correlation between temp and kWh in\nthis sample.')
    })
```

## Second plot

The consumption profile becomes more predictable as the sample size increases.  Here is an example: 3 distinct households have very different profiles, but as the sample size increases to 100 the aggregate consumption converges to the lower profile.

```{r echo=FALSE, message=FALSE}
l1 <- plot_ly(df_13Feb, x = ~interval, y = ~N0191, type = 'scatter', mode = 'lines', height = 150) %>%
  layout(xaxis = list(title = '', showticklabels = FALSE), yaxis = list(title = 'kWh'))
l2 <- plot_ly(df_13Feb, x = ~interval, y = ~N0177, type = 'scatter', mode = 'lines', height = 150) %>%
  layout(xaxis = list(title = '', showticklabels = FALSE), yaxis = list(title = 'kWh'))
l3 <- plot_ly(df_13Feb, x = ~interval, y = ~N0240, type = 'scatter', mode = 'lines', height = 150) %>%
  layout(xaxis = list(title = '', showticklabels = FALSE), yaxis = list(title = 'kWh'))

l <- subplot(l1, l2, l3, nrows = 1, shareX = TRUE, shareY = TRUE)

l4 <- plot_ly(tempdf_13Feb, x = ~interval, y = ~kWh, type = 'scatter', mode = 'lines', height = 350)

subplot(l, l4, nrows = 2, heights = c(0.3, 0.7), titleX = TRUE, titleY = TRUE) %>%
  layout(showlegend = FALSE)
```
