---
title: "Electricity Consumption and Temperature"
author: "Michael Tyler"
date: "21 November 2017"
output: ioslides_presentation
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```

## Synopsis

- Here we investigate whether there is a correlation between electricity consumption and outdoor temperature for a single London household equipped with a 'smart' electricity meter.
- The electricity consumption data is a sample from the Low Carbon London trial, in which ~5000 London homes were fitted with smart meters and their consumption recorded at half-hourly intervals for a period of a year, between 2012 and 2013.
- The weather data is a random sample of 30 days during this year, obtained from the Weather Underground website.

## First plot

```{r echo=FALSE, message=FALSE}
library(lubridate)
library(jsonlite)
library(dplyr)
library(data.table)
library(plotly)
```

Here is a plot of electricity consumption per half-hour interval (in kWh) against temperature (in Celsius).

```{r echo=FALSE, message=FALSE}
df <- read.csv('kWh_temp.csv')

mod <- lm(kWh ~ temp, df)

p <- plot_ly(df, x = ~temp, y = ~kWh, type = 'scatter')
add_lines(p, x = ~temp, y ~ fitted(mod))
```

## Coefficients of first model

The linear model shows a significant negative correlation between electricity use and temperature:

```{r echo=FALSE}
summary(mod)$coeff
```

## Second plot

The values above 0.4kWh appear to have high influence on the linear model.  Here is a plot with these values removed.

```{r echo=FALSE, message=FALSE}
df2 <- df[df$kWh <= 0.4, ]

mod2 <- lm(kWh ~ temp, df2)

p2 <- plot_ly(df2, x = ~temp, y = ~kWh, type = 'scatter')
add_lines(p2, x = ~temp, y ~ fitted(mod2))
```

## Coefficients of second model

The new linear model still shows a significant negative correlation between electricity use and temperature:

```{r echo=FALSE}
summary(mod2)$coeff
```

## Thanks,

and I hope that was interesting.

